---
title: "p8105_hw3_TZ2472"
author: "tianwei zhao"
date: "10/12/2021"
output: github_document
---

```{r}
library(tidyverse)


theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## Problem 1


The goal is to do some exploration of this dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations. Then, do or answer the following (commenting on the results of each):

```{r}
library(p8105.datasets)
data("instacart")


#an illustrative example of the first order:
instacart %>% 
  filter(order_id%in%  "1")

```

In the instacart data set, we have a total of `r nrow(instacart)` observations, each represents an order in instacart .The data has `r ncol(instacart)` variables, collects key information about instacart such as : order identifier, product identifier, order in which each product was added to cart, reordered records, customer identifier, the order sequence number for usesr, the day of the week on which the order was placed, the hour of the day on which the order was placed, days since the last order, name of the product, aisle identifier,department identifier, name of the aisle,  name of the department.

Take example of the first order of instacart,On Thursday 10 o'clock, an user ID 112108 check out 8 items that are Bulgarian Yogurt, Organic 4% Milk Fat Whole Milk Cottage Cheese, Organic Celery Hearts,	Cucumber Kirby,	Lightly Smoked Sardines in Olive Oil,	Bag of Organic Bananas,	Organic Hass Avocado, Organic Whole String Cheese(exactly in this order).Those items are in aisle of yogurt,other creams cheeses,fresh vegetables,fresh vegetables,canned meat seafood,fresh fruits,	fresh fruits,and packaged cheese.It has been 9 days since the user 112108's last order.Among the orders, Bulgarian Yogurt, Organic 4% Milk Fat Whole Milk Cottage Cheese,Lightly Smoked Sardines in Olive Oil,Organic Whole String Cheese are ordered before, and the rest are new tries to this user.



How many aisles are there, and which aisles are the most items ordered from?

```{r}
instacart %>% 
  count(aisle,aisle_id) %>% 
  arrange(desc(n))

```

The result data set(134*3) gives us the most ordered items in the first row,from the table we can find there are 134 aisles and the fresh vegetables is the most items ordered from.



Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r}
  instacart %>% 
  count(aisle) %>% 
  arrange(n) %>% 
  filter(n>10000) %>% 
  mutate(
    aisle = factor(aisle),
    aisle = forcats::fct_reorder(aisle, sort(n))) %>% 
  ggplot(aes(x=n,y=aisle,color = n))+
  geom_point()+
  labs(
    title = "Number of items ordered in Insta's aisles plot",
    x = "Number of items ordered",
    y = "Aisle",
    caption = "Data from the instacart"
  )+
  scale_x_continuous(
    breaks = c(0,20000,40000,60000,80000,100000,120000,140000)
  )

```

I arrange the aisles in a increasing order of number of items ordered for better readability, and use color function to denote the difference of number of items ordered.The lighter the color is , the more the item ordered.



Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
new_df = instacart %>% 
  filter(aisle%in%c("baking ingredients", "dog food care","packaged vegetables fruits")) %>% 
  group_by(product_name,aisle) %>% 
  summarize(n_obs=n())

baking_df = 
  filter(new_df,aisle%in% "baking ingredients") %>% 
  arrange(desc(n_obs)) %>% 
  filter(n_obs> 330)

dog_df = 
  filter(new_df,aisle%in% "dog food care") %>% 
  arrange(desc(n_obs)) %>% 
  filter(n_obs> 25)

vegfruit_df = 
  filter(new_df,aisle%in% "packaged vegetables fruits") %>% 
  arrange(desc(n_obs)) %>% 
  filter(n_obs> 4060)

top3_table = 
  bind_rows(baking_df,dog_df,vegfruit_df) %>% 
  relocate(aisle) %>% 
  rename(num_of_time_ordered = n_obs) %>% 
  knitr::kable(caption = "TOP3 Popular Items")

top3_table

```

In the table top3_table, it contains the top 3 popular items' names and number of times ordered in baking ingredients aisle, dog food care aisle, and packaged vegetables fruits aisle.
 
 

Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r}
pinkapple_df = instacart %>% 
  filter(product_name%in%c("Pink Lady Apples","Coffee Ice Cream"))%>% 
  group_by(order_dow,product_name) %>% 
  summarize(
    mean_hour = mean(order_hour_of_day)) %>% 
  pivot_wider(
    names_from = "order_dow",
    values_from = "mean_hour",
  ) %>% 
  rename("Sunday" = "0", 
         "Monday" = "1", 
         "Tuesday" = "2", 
         "Wednesday" = "3",
         "Thursday" = "4",
         "Friday" = "5",
         "Saturday" = "6") %>% 
  knitr::kable(caption = "Mean Hour on Each Day of the Week")
  
pinkapple_df
```

This table shows the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. The mean hour of the day that people order Pink Lady Apples and Coffee Ice Cream is around 11 o'clock and 14 o'clock. The mean varies very slightly in each day of the week. 


##Problem 2 


```{r}
library(p8105.datasets)
data("brfss_smart2010")
```

First, do some data cleaning:

format the data to use appropriate variable names;
focus on the “Overall Health” topic
include only responses from “Excellent” to “Poor”
organize responses as a factor taking levels ordered from “Poor” to “Excellent”

```{r}
mydata_df=brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  rename(
    state = locationabbr,
    location = locationdesc
  ) %>% 
  filter(topic%in%"Overall Health",
         response%in% c("Excellent","Very good","Good","Fair","Poor")) %>% 
  relocate(topic,response) %>% 
  mutate(
    response = factor(response),
    response = forcats::fct_relevel(response,c("Poor","Fair","Good","Very good","Excellent"))
  ) %>% 
  arrange(response)
mydata_df
```




Using this dataset, do or answer the following (commenting on the results of each):

In 2002, which states were observed at 7 or more locations? What about in 2010?

```{r}
mydata_df %>% 
  filter(year == 2002) %>% 
  group_by(state) %>%
  summarize(n_obs=n_distinct(location)) %>% 
  filter(n_obs>=7) %>% 
  select(state)

mydata_df %>% 
  filter(year == 2010) %>% 
  group_by(state) %>%
  summarize(n_obs=n_distinct(location)) %>% 
  filter(n_obs>=7) %>% 
  select(state)


```

In 2002, there are 6 states that were observed at 7 or more locations. Those states are CT,FL,MA, NC, NJ, and PA. 

In 2010, there are 14 states that were observed at 7 or more locations. Those states are CA, CO, FL, MA, MD, NC, NE, NJ, NY, OH, PA, SC, TX, and WA. 


Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).

```{r}
mydata_df %>% 
  filter(response == "Excellent") %>% 
  group_by(year,state) %>% 
  mutate(mean_data_value = mean(data_value,rm.na=TRUE)) %>% 
  select(year,state,mean_data_value) %>% 
  ggplot(aes(x = year,y = mean_data_value,color = state))+
  geom_line(aes(group = state))+
  theme(legend.position = "right")
  
  
```

Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.

```{r}
mydata_df %>% 
  filter(year%in%c(2006,2009),
         state == "NY") %>% 
  ggplot(aes(x = data_value, fill = response)) +
  geom_density(alpha = .5) + 
  facet_grid(. ~ year) +
  theme(legend.position = "bottom")
```


## Problem 3

Load, tidy, and otherwise wrangle the data. Your final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes. Describe the resulting dataset (e.g. what variables exist, how many observations, etc).

```{r}
accel_df = read_csv("accel_data.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
   "activity_1":"activity_1440",
    names_to = "activity_order",
    names_prefix = "activity_",
    values_to = "activity") %>% 
  rename(activity_counts = activity) %>% 
  mutate(
    day_type = if_else(day %in% c("Saturday","Sunday"), "weekend","weekday"),
    week = as.integer(week),
    day_id = as.integer(day_id),
    activity_order = as.integer(activity_order)
          ) %>% 
  select(week,day_id,day, day_type,activity_order,activity_counts)

```

This data set have `r ncol(accel_df)` variables, including the week, day, day_id, minute of the day, day_type, and activity value. The set has `r nrow(accel_df)` observations. Each observation is about the activity counts for a specific minute from a 24-hour day and the information about that day, including the week, day_id, and whether it is a weekday or weekend.


Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?

```{r}
accel_df %>% 
  group_by(day_id) %>% 
  mutate(total_activity = sum(activity_counts)) %>% 
  select(day_id, day, day_type, total_activity) %>% 
  distinct() %>% 
  knitr::kable()

```

No, there is no trends apparent.


Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.

```{r}
total_activity_df = accel_df %>% 
  group_by(day_id) %>% 
mutate(
    total_activity = sum(activity_counts),
    hour = as.numeric(activity_order)/60)

total_activity_df %>% 
  group_by(week) %>% 
  ggplot(aes(x = hour, y = activity_counts, color = day)) +
  geom_smooth()
  
```

From the plot we can see that the activity time for hours differs significantly from one day to another. Moreover, even for the same day of the week, the activity time courses are very different week from week. This is abnormal, suggesting that there might be some mistake in the original data, especially the recording in Saturday, where the participants had as high as  607175 counts in week two and as low as 1440 in week 4 and 5. Therefore, we might have to go back to the original data and correct the data before further analysis.
